PhD Thesis

### 2016-12-04 thesis-main.tex
Misc. Revise

However, often queries are executed on collections of data that are not stored
in a database, but in collections manipulated by some program. Moving in-memory
data to RDBMSs, typically does not improve
performance~\citep{Stonebraker07,Rompf2015functional}, and reusing database
optimizers is not trivial.

Moreover, many programming languages are far more expressive than RBDMSs.
Typical RDBMS can only manipulate SQL relations, that is multisets (or bags) of
tuples (or sometimes sets of tuples, after duplicate elimination). Typical
programming languages (PL) support also arbitrarily nested lists and maps of
data, and allow programmers to define new data types with few restrictions; a
typical PL will also allow a far richer set of operations than SQL.

@@

We consider the problem for functional programming languages such as Haskell or
Scala, and we consider collection queries written using the APIs of their
collection libraries, which we treat as an embedded domain-specific language (EDSL). Such
APIs (or DSLs) contain powerful operators on collections such as \texttt{map}\pg{update
  style to be uniform with the one from lhs2TeX}, which are higher-order, that
is they take as arguments arbitrary functions in the host language that we must
also handle.

@@

In particular, this thesis is divided into two parts:
\begin{itemize}
\item we first describe, in Chapter 2, work on optimizing collection queries by static program transformation.
\item in the remaining chapters, we describe work on incrementalizing programs
  by static program transformation. This thesis presents the first approach that
  handles higher-order programs by using program transformation; hence, while
  our main examples use collection queries, we phrase the work in terms of
  $\lambda$-calculi with unspecified primitives.
\end{itemize}
% Modern libraries offer high-level APIs for manipulating (in-memory) collections;
% equivalently, such an API is an embedded domain-specific language (EDSL).

@@

\chapter{Reifying collection queries}

>>>>

However, often queries are executed on collections of data that are not stored
in a database, but in collections manipulated by some program. Moving in-memory
data to RDBMSs typically does not improve
performance~\citep{Stonebraker07,Rompf2015functional}, and reusing database
optimizers is not trivial.

Moreover, many programming languages are far more expressive than RDBMSs.
Typical RDBMS can only manipulate SQL relations, that is multisets (or bags) of
tuples (or sometimes sets of tuples, after duplicate elimination). Typical
programming languages (PL) support also arbitrarily nested lists and maps of
data, and allow programmers to define new data types with few restrictions; a
typical PL will also allow a far richer set of operations than SQL.

@@

We consider the problem for functional programming languages such as Haskell or
Scala, and we consider collection queries written using the APIs of their
collection libraries, which we treat as an embedded domain-specific language (EDSL). Such
APIs (or DSLs) contain powerful operators on collections such as $\Varid{map}$, which are higher-order, that
is they take as arguments arbitrary functions in the host language that we must
also handle.

@@

In particular, this thesis is divided into two parts:
\begin{itemize}
\item we first describe, in \cref{sec:ch-aosd13}, work on optimizing collection queries by static program transformation.
\item in the remaining chapters, we describe work on incrementalizing programs
  by static program transformation. This thesis presents the first approach that
  handles higher-order programs by using program transformation; hence, while
  our main examples use collection queries, we phrase the work in terms of
  $\lambda$-calculi with unspecified primitives.
\end{itemize}
% Modern libraries offer high-level APIs for manipulating (in-memory) collections;
% equivalently, such an API is an embedded domain-specific language (EDSL).

@@

\chapter{Reifying collection queries}
\label{sec:ch-aosd13}
